Economics and Crime Patterns: The Interplay Between CPI Changes and Vehicle Theft Frequencies

### 2024-04-19 paper/paper.qmd
update 3.0 version of paper

---
title: "Economics and Crime Patterns: The Interplay Between CPI Changes and Vehicle Theft Frequencies"
subtitle: "Insights from Predictive Analytics and Visual Data Exploration"
author: 
  - Siqi Fei
thanks: "Code and data are available at: <https://github.com/FXXFERMI/Inflation-and-Crime-Patterns.git>. Link to the Shiny App:  <https://siqi-fei.shinyapps.io/VehicleTheft_Inflation/>"
date: today
date-format: long
abstract: "In this study, we researched the monthly incidence of vehicle thefts and inflation from 2014 to 2024 by using data sourced from the Toronto Police Service and Statistics Canada. We analyzed the data distribution and characteristics of the two datasets and combined both to conduct a correlation analysis by using Bayesian multiple linear regression model and a Shiny application. We found a significant positive correlation between the overall CPI and vehicle theft rates.  This finding suggests that higher inflation may be associated with an increased number of vehicle theft. This study highlights the importance of economic indicators in predicting criminal behaviour which in terms provide valuable insights for policymakers and economic analysts."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

@@

@@

```{r}
#| label: tbl-analysisData
#| tbl-cap: Sample of Analysis Dataset
#| echo: false

selected_data <- cleaned_data[, c(1, 7, 3, 6, ncol(cleaned_data))]
kable(head(selected_data), caption = "Sample of Analysis Dataset")
```

## Data Distribution

@@

::: {#fig-CPIDistribution}
![](images/box_all_cpi.png){width="73%"}

@@

## Model set-up

@@

## Model justification

@@

::: {#fig-TimeSeries}
![](images/Time.png)

@@

The @tbl-CPIData shows a sample dataset of CPI from @statscanCPI2023. The @tbl-AutoData shows a sample dataset of Vehicle Thetfs Dataset from @torontoPoliceAutoTheft2023.

>>>>

---
title: "Economics and Crime Patterns: The Interplay Between CPI Changes and Vehicle Theft Frequencies"
subtitle: "Insights from Predictive Analytics and Visual Data Exploration"
author: 
  - Siqi Fei
thanks: "Code and data are available at: <https://github.com/FXXFERMI/Inflation-and-Crime-Patterns.git>. Link to the Shiny App:  <https://siqi-fei.shinyapps.io/VehicleTheft_Inflation/>"
date: today
date-format: long
abstract: "In this study, we researched the monthly incidence of vehicle thefts and inflation from 2014 to 2024 by using data sourced from the Toronto Police Service and Statistics Canada. We analyzed the data distribution and characteristics of the two datasets and combined both to conduct a correlation analysis by using Bayesian multiple linear regression model and a Shiny application. We found a significant positive correlation between the overall CPI and vehicle theft rates.  This finding suggests that higher inflation may be associated with an increased number of vehicle theft. This study highlights the importance of economic indicators in predicting criminal behaviour which in terms provide valuable insights for policymakers and economic analysts."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
frontsize: 10pt
---

@@

## Data Cleaning

@@

```{r}
#| label: tbl-analysisData
#| tbl-cap: Sample of Analysis Dataset (includes only three CPI categories)
#| echo: false

selected_data <- cleaned_data[, c(1, 7, 3, 6, ncol(cleaned_data))]
kable(head(selected_data), caption = "Sample of Analysis Dataset (includes only three CPI categories)")
```

@@

::: {#fig-CPIDistribution}
![](images/box_all_cpi.png){width="75%"}

@@

## Model Set-up

@@

## Model Justification

@@

::: {#fig-TimeSeries}
![](images/Time.png){width="110%"}

@@

The @tbl-CPIData shows a sample dataset of CPI from @statscanCPI2023, it shows all the CPI categories in the dataset, while only have 3 observations. The @tbl-AutoData shows a sample dataset of Vehicle Thetfs Dataset from @torontoPoliceAutoTheft2023.
